### 2022-08-25 chapters/optimization/java-dir/java.tex
add : bitops analysis

\noindent\textbf{Software Settings.}
For the sake of reproducibility, each experiment runs within a Docker container based on \textsf{SDKMAN!}\footnote{\url{https://sdkman.io}} image and Alpine docker.\footnote{\url{https://github.com/alpinelinux/docker-alpine}}

@@

\vspace{6pt}
\noindent\textbf{Java Virtual Machines Candidates.}
We considered a set of $52$ JVM distributions taken from $8$ different providers/packagers mostly obtained from \textsf{SDKMAN!}, as listed in \Cref{tab:JVM}.
Depending on providers, either all the versions, majors, or LTS are made available by \textsf{SDKMAN!}.

@@

\subsection{Extension}
We also added an extension to the protocol to allow the user to run the same experiment with different configurations.
The package is available in the GitHub repository.\footnote{\url{https://github.com/chakib-belgaid/jvm-comparaison}}
To add extra \textbf{jvm candidates} for the benchmark applications, we added a new configuration file \texttt{jvm.sh} in the root directory of the repository, where we put the name and the version of the jvm to be used.\note{The JVM must be supported by \textsf{SDKMAN!}}
For the \textbf{input workload},the benchmarks should be provided in the \texttt{benchmarks} directory.
As for extra \textbf{metrics}, one can create a new script file that monitors the experiment and record the metrics.
We provide some examples, such as \texttt{recordpower.sh} to measure the instantaneous power and \texttt{recordthreads.sh} to measure the number of active threads during the experiment.

>>>>

\noindent\textbf{Software Settings.}
For the sake of reproducibility, each experiment runs within a Docker container based on \textsf{SDKMAN}\footnote{\url{https://sdkman.io}} image and Alpine docker.\footnote{\url{https://github.com/alpinelinux/docker-alpine}}

@@

\vspace{6pt}
\noindent\textbf{Java Virtual Machines Candidates.}
We considered a set of $52$ JVM distributions taken from $8$ different providers/packagers mostly obtained from \textsf{SDKMAN}, as listed in \Cref{tab:JVM}.
Depending on providers, either all the versions, majors, or LTS are made available by \textsf{SDKMAN}.

@@

\subsection{Extension}
We also added an extension to the protocol to allow the user to run the same experiment with different configurations.
The package is available in the GitHub repository.\footnote{\url{https://github.com/chakib-belgaid/jvm-comparaison}}
To add extra \textbf{jvm candidates} for the benchmark applications, we added a new configuration file \texttt{jvm.sh} in the root directory of the repository, where we put the name and the version of the jvm to be used.
For the \textbf{input workload},the benchmarks should be provided in the \texttt{benchmarks} directory.
As for extra \textbf{metrics}, one can create a new script file that monitors the experiment and record the metrics.
We provide some examples, such as \texttt{recordpower.sh} to measure the instantaneous power and \texttt{recordthreads.sh} to measure the number of active threads during the experiment.

### 2022-08-25 chapters/optimization/python-dir/python_interperters.tex
refactoring
check
add : bitops analysis

\subsection{Preliminary stduies}
This part will be dedicated to analyse the initial behaviour of different python interpereters.
First we started with a typical benchmark \em binarrytree. where we compare the energy consumption of the different implementations.

\section{placeholder}

we performed a shapiro normality test for the first on the results, and almost all the p-values smaller than alpha=0.01\% therefore the distrubution is not normal

@@

\begin{figure}

    \label{table:bitops}
    \begin{tabular}{|lrrrrr|}
        \toprule
        benchmark    & A       & GG      & LL      & Or      & XOR     \\
        \midrule
        activepython & 676.980 & 763.208 & 651.783 & 743.016 & 728.828 \\
        cpython2     & 441.082 & 435.886 & 430.846 & 415.247 & 419.081 \\
        cpython3     & 595.209 & 685.085 & 563.839 & 657.972 & 655.560 \\
        cython       & 35.077  & 182.688 & 274.177 & 34.868  & 34.504  \\
        nuitka       & 33.260  & 32.980  & 33.256  & 33.472  & 33.030  \\
        numba2       & 9.102   & 8.411   & 9.460   & 9.375   & 9.755   \\
        numba3       & 9.566   & 10.144  & 9.219   & 9.344   & 9.665   \\
        pypy2        & 8.456   & 7.844   & 8.286   & 8.138   & 7.952   \\
        pypy3        & 7.552   & 8.093   & 8.108   & 8.669   & 8.623   \\
        shedskin     & 8.024   & 8.070   & 8.399   & 8.126   & 8.277   \\
        \bottomrule
    \end{tabular}
    \vfill
    \centering
    \includegraphics[width=\linewidth]{imgs/bitopts_mean}
    \caption{energy consumption of different impelementations using Bit Operation benchmarks (Joule) }
\end{figure}

@@

\begin{figure}

\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/dendogram_interpreters}
    \caption{Dendogram of the different implementations}
    \label{fig:dendogram}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/alltomti_performance}
    \caption{different interpererts optimisation }
    \label{fig:tommi_all}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__pypy2_pypy3}
    \caption{green factor of pypy }
    \label{fig:pypy2vspypy3}
\end{figure}

@@

As we can see in figure \ref{fig:tommi_all}, there is no evolution between cpython2 and cpython3
Intelpython and active python both folows the same behaviour. one can conclude that the work that have been on those interpreters is mainly
to improve a specific purpose, Active python claims that their version is focused on secuirty which explains the lack of some performances due to the introduction of more reonforcment
.Morver Intel published their version of python as a dedicated for machine learning. Unfortunately the tommti benchmark is a set that focuses on the general purpose programming
which does not reflect the performance of the machine learning.
Another aspect of a such behaviour might be due to the fact of the processors that were used in the testbed. and it might change in the future if we include the GPU part for the machine learning benchmarks
% reference the work of the intern 
As for nuitka. there were no optimization in the energy consumpotion despite the fact that is it a complier.
However if we dig through the nuitka mechanisms. they basically embed the python code with an interpreter.
Unllike nuitka shedskin exhebit a the best energy consumption pattern when it comes to the arithmetic operartions. One can conclude it is due to the fact of the native type of the variables. unlike the interpreters where they are treated as object in the begining.

@@

\input{table-envs.tex}

>>>>

\subsection{Preliminary stduies}
This part will be dedicated to analyse the initial behaviour of different python interpereters.
First we started with a typical benchmark binarrytree. where we compare the energy consumption of the different implementations.

\subsection{python and binarry operations }
The goal of this study is to look at how different Python implementations work while treating different binary operations .

\begin{figure}[!hbt]
    \begin{tabular}{|lrrrrr|c|}
        \toprule
        benchmark    & Addition & Right Rotation & Left Rotation & Or      & XOR     & Average \\
        \midrule
        activepython & 676.980  & 763.208        & 651.783       & 743.016 & 728.828 & 712.76  \\
        cpython2     & 441.082  & 435.886        & 430.846       & 415.247 & 419.081 & 428.43  \\
        cpython3     & 595.209  & 685.085        & 563.839       & 657.972 & 655.560 & 531.53  \\
        \hline
        cython       & 35.077   & 182.688        & 274.177       & 34.868  & 34.504  & 112.26  \\
        nuitka       & 33.260   & 32.980         & 33.256        & 33.472  & 33.030  & 33.2    \\
        \hline
        numba2       & 9.102    & 8.411          & 9.460         & 9.375   & 9.755   & 9.22    \\
        numba3       & 9.566    & 10.144         & 9.219         & 9.344   & 9.665   & 9.59    \\
        pypy2        & 8.456    & 7.844          & 8.286         & 8.138   & 7.952   & 8.13    \\
        pypy3        & 7.552    & 8.093          & 8.108         & 8.669   & 8.623   & 8.21    \\
        shedskin     & 8.024    & 8.070          & 8.399         & 8.126   & 8.277   & 8.18    \\
        \bottomrule
    \end{tabular}
    % should i add the interpretation 
    \caption{energy consumption of different impelementations using Bit Operation benchmarks (Joule) }
    \label{table:bitops}
\end{figure}

Table~ provides a detailed report on the energy consumption of different implementations of the code.

Except for \emph{Cython}, each implementation treats the binary operation the same. As one can see, Table~\ref{table:bitops} shows no difference in the energy consumption of those operations.

However, when it comes to Cython, the rotation operations tend to consume 5 times more energy than the other operations, and left rotations consume even more than the right ones.

The reason behind such behavior is the fact that Cython does not have a native implementation for these two operations, so it replaces them with more complex operations, which leads to an extra time of 6s for the right rotation and 10s for the left, compared to 1.5 for the other operations. -

Therefore, it is safe to summarize all the operations into one that we call binary.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/bitopts_mean}
    \caption{Summary of the binary operations for different python implementations }
    \label{figure:bitops}
\end{figure}
Figure \ref{figure:bitops}  shows the average energy consumption of different implementations for the binary operations.

One can notice 3 clusters.

The first, which is the interpreted one, uses approximately 500 joules per operation. The second is a non-optimized compilation that includes Cython and Nuitka and uses 10 times less energy than the first. And the final one, which has been fully optimized, uses 50 percent less energy. This category includes jit libraries, also known as numba, Python implementations that include JIT, also known as pypy, and shedskin, which transpile the source code into C and then use a C/C++ compiler to generate the binary, as opposed to cython and nuitka, which directly compile the python code.

To conclude this. JIT has a huge impact when it comes to binary functions, and the wrong choice of the implementation can lead to up to a 100x increase in energy consumption.

\section{placeholder}
introdution to the stufy
we performed a shapiro normality test for the first on the results, and almost all the p-values smaller than alpha=0.01\% therefore the distrubution is not normal

@@

@@

\input{table-envs.tex}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/dendogram_interpreters}
    \caption{Dendogram of the different implementations}
    \label{fig:dendogram}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/alltomti_performance}
    \caption{different interpererts optimisation }
    \label{fig:tommi_all}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__pypy2_pypy3}
    \caption{green factor of pypy }
    \label{fig:pypy2vspypy3}
\end{figure}

@@

%   this is after , when we come to the conclusion
As we can see in figure \ref{fig:tommi_all}, there is no evolution between cpython2 and cpython3
Intelpython and active python both folows the same behaviour. one can conclude that the work that have been on those interpreters is mainly
to improve a specific purpose, Active python claims that their version is focused on secuirty which explains the lack of some performances due to the introduction of more reonforcment
.Morver Intel published their version of python as a dedicated for machine learning. Unfortunately the tommti benchmark is a set that focuses on the general purpose programming
which does not reflect the performance of the machine learning.
Another aspect of a such behaviour might be due to the fact of the processors that were used in the testbed. and it might change in the future if we include the GPU part for the machine learning benchmarks
% reference the work of the intern
As for nuitka. there were no optimization in the energy consumpotion despite the fact that is it a complier.
However if we dig through the nuitka mechanisms. they basically embed the python code with an interpreter.
Unllike nuitka shedskin exhebit a the best energy consumption pattern when it comes to the arithmetic operartions. One can conclude it is due to the fact of the native type of the variables. unlike the interpreters where they are treated as object in the begining.

@@
